EinsteinPy: Python for General Relativity\footnote{Released on July, 15th, 2019}

### 2020-01-20 paper.tex
completed schwarzschild

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.3
\documentclass{aastex63}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}

@@

\section{The EinsteinPy Core Package}
This section explains in detail, the modules available within EinsteinPy core pacakge, their API and the purpose they serve.
\subsection{METRIC}
This module captures all the geometric and causal structure of specific spacetimes, and being able to calculate trajectories in a given space-time. 
The module derives its name from Metric Tensor, a tensorial quantity used to describe differential lengths, especially in a curved space-time. 
However, this module serves more than its usual definition and ultimately leads us to obtain geodesics of particles in a spacetime with high curvature, where notions of newtonian physics fail.
\subsubsection{Schwarzschild Metric}
Karl Schwarzchild pioneered the analysis of the relation between the size of black hole and its mass and this work paved the way for the first exact solution of Einstein's Field Equation under the limited case of single spherical non-rotating, non-charged body. 
The metric equation corrosponding to a space-time centred around a body in the above described conditions is given by,

\begin{align}
& ds^2 = -d\tau ^2 = - (1 - \frac{r_s}{r})c^2 dt^2  + \frac{1}{(1 - \frac{r_s}{r})}dr^2 + r^2 d\theta ^2 + r^2 sin^2 \theta d\phi ^2
\end{align}

@@

>>>>

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.3
\documentclass{aastex63}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{listings}  % for code in latex
\usepackage{color}

@@

\section{The EinsteinPy Core Package} \label{sec:CORE}
This section explains in detail, the modules available within EinsteinPy core pacakge, their API and the purpose they serve.
\subsection{METRIC}
This module captures all the geometric and causal structure of specific spacetimes, and being able to calculate trajectories in a given space-time. 
The module derives its name from Metric Tensor, a tensorial quantity used to describe differential lengths, especially in a curved space-time. 
However, this module serves more than its usual definition and ultimately leads us to obtain geodesics of particles in a spacetime with high curvature, where notions of newtonian physics fail.
\subsubsection{Schwarzschild Metric}
Karl Schwarzchild pioneered the analysis of the relation between the size of black hole and its mass and this work paved the way for the first exact solution of Einstein's Field Equation under the limited case of single spherical non-rotating, non-charged body. 
The metric equation corrosponding to a space-time centred around a body in the above described conditions is given by,

\begin{align}
& ds^2 = -c^2 d\tau ^2 = - (1 - \frac{r_s}{r})c^2 dt^2  + \frac{1}{(1 - \frac{r_s}{r})}dr^2 + r^2 d\theta ^2 + r^2 sin^2 \theta d\phi ^2
\end{align}

@@

Schwarzschild Radius is a limit below which the degree of space-time curvature causes the partcle to undergo irreversible gravitational collapse wherein the partcle would need a speed greater than speed of light ot escape singularity.

As $r \to \infty$, the metric defined above starts to approximate flat Minkowski space-time, thus showing that Schwarzschild geometry is asymptotically flat.

If supplied with initial position and velocity of a particle with negligible mass(in comparision to the central body), we can obtain the geodesics of the particle for a given range of proper time $\tau$, as it boils down to a set of eight differential equations numerically solvable by a class of Runge-Kutta methods.

The file \textbf{schwarzschild.py} contains the class \textit{Schwarzschild} for defining Schwarzschild geometry methods.

Here is a working code describing the motion of particle revolving around a black hole of mass approximately equal to mass of the Earth. As the trajectory starts from a mere $130m$ from the black hole, we obtain some interesting observations.

\begin{lstlisting}[language=Python, caption=Importing the required modules]
>>> import numpy as np
>>> import astropy.units as u
>>> from einsteinpy.plotting import GeodesicPlotter
>>> from einsteinpy.coordinates import SphericalDifferential
>>> from einsteinpy.bodies import Body
>>> from einsteinpy.geodesic import Geodesic
>>> from einsteinpy.metric import Schwarzschild
\end{lstlisting}

\begin{lstlisting}[language=Python, caption=Defining the attractor and the particle and calculating geodesics]
>>> Attractor = Body(name="BH", mass=6e24 * u.kg)
>>> init_conditions = SphericalDifferential(130*u.m, np.pi/2*u.rad, -np.pi/8*u.rad,
...                                         0*u.m/u.s, 0*u.rad/u.s, 1900*u.rad/u.s)
>>> Particle = Body(differential=init_conditions, parent=Attractor)
>>> geodesic = Geodesic(body=Particle, time=0 * u.s, end_lambda=0.00205, step_size=5e-8)
>>> goedesic.trajectory
array([[ 0.00000000e+00,  1.20104339e+02, -4.97488462e+01, ...,
         9.45228078e+04,  2.28198245e+05,  0.00000000e+00],
       [ 3.99986427e-08,  1.20108103e+02, -4.97397110e+01, ...,
         9.36472607e+04,  2.28560869e+05, -5.80280784e-14],
       ...,
       [ 2.04962579e-03,  1.04114904e+02,  1.08255773e+01, ...,
        -1.23245830e+06,  1.80261847e+05, -9.67931265e-09],
       [ 2.05002578e-03,  1.03619005e+02,  1.08973776e+01, ...,
        -1.24706334e+06,  1.78734562e+05, -9.79337222e-09]])
>>> geodesic.trajectory.shape
(6567, 8)
\end{lstlisting}
We can see that for each timestep, we obtain eight values : $t$, $r$, $\theta$, $\phi$, $v_t$, $v_r$, $v_\theta$ and $v_\phi$ corrosponding to coordinate time, radial distance from the centre, polar angle, azimuth angle and their respective velocities w.r.t proper time $\tau$.

Further, we plot the trajectory in 2D for vizualization purposes.
\begin{lstlisting}[language=Python, caption=Plotting the trajectory]
>>> obj = GeodesicPlotter()
>>> obj.plot(geodesic)
>>> obj.show()
\end{lstlisting}

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.7]{images/perihelion_cropped.png}
	%\centerline{\includegraphics{ftcbuild2.pdf}}
	\caption{Plot obtained from the above code showing the phenomenon of perihelion advance}
	\label{fig:schwarzschild}
\end{figure}

We observe the phenomenon of perihelion advance of a particle around a heavy body due to spacetime curavture. The orbit would have been purely eliptical if plotted in accordance with Newtonian Mechanics, however, the curvature of space-time around the heavy mass shifts the orbit with each revolution. In practicality, this phenomenon is not so pronounced as no particle approaches a heavy mass this closely. For instance, Mercury advances its perihelion at a rate of $42.98\pm 0.04$ arcsec/century.
